Predicting Probability of Participants' Vote of Conservative and Liberty in 2025 Canadian election 

### 2024-04-14 paper/paper(1).qmd
Add files via upload

Canadian Election Study gathered a sample of 20,968 individuals from across Canada using the Leger Opinion panel, ensuring representation from various regions while maintaining a balance in gender and age demographics within each area.
The sample aimed for an equal distribution of 50% males and 50% females, with age groups targeted as follows: 28% between 18-34 years, 33% between 35-54 years, and 39% 55 years and older.
The geographical regions included Atlantic Canada, Quebec, Ontario, Western Canada, and the Territories, with quotas set to reflect their population proportions (7% for Atlantic, 23% for Quebec, 38% for Ontario, and 32% for the West) and an open acceptance for any participants from the Territories.
In Quebec, the language preference aimed for 80% French-speaking and 20% English-speaking participants, with a goal of 10% French-speaking participants in both the Atlantic region and across Canada.
Eligibility for participation required being at least 18 years old and a Canadian citizen or permanent resident[@CES].

The General Social Survey focuses on non-institutionalized individuals aged 15 and above residing in Canada's ten provinces.
Within each selected household, a random individual is chosen by an application to answer the survey following the initial household roster completion.
General Social Survey's questionnaire was developed through rigorous research and broad consultations with data users.
The Statistics Canada Questionnaire Design Resource Center (QDRC) conducted qualitative testing in four cities to identify effective questions and those requiring adjustments.
Based on the testing, QDRC prepared a comprehensive report with suggestions, which were integrated into the final survey design to enhance clarity and effectiveness.The survey employs a cross-sectional sample survey methodology, utilizing a combined frame of landline and cellular phone numbers from Census data, administrative records, and Statistics Canada's dwelling database.
This approach improves coverage of households linked to phone numbers.
The sampling strategy is stratified, with stratification occurring at the level of provinces and census metropolitan areas (CMA), and it relies on probability sampling techniques.
Information is collected from a single individual aged 15 or older per household, without the use of proxy responses.[@GSS]

@@

# Discussion

>>>>

Canadian Election Study gathered a sample of 20,968 individuals from across Canada using the Leger Opinion panel, ensuring representation from various regions while maintaining a balance in gender and age demographics within each area.
The sample aimed for an equal distribution of 50% males and 50% females, with age groups targeted as follows: 28% between 18-34 years, 33% between 35-54 years, and 39% 55 years and older.
The geographical regions included Atlantic Canada, Quebec, Ontario, Western Canada, and the Territories, with quotas set to reflect their population proportions (7% for Atlantic, 23% for Quebec, 38% for Ontario, and 32% for the West) and an open acceptance for any participants from the Territories.
In Quebec, the language preference aimed for 80% French-speaking and 20% English-speaking participants, with a goal of 10% French-speaking participants in both the Atlantic region and across Canada.
Eligibility for participation required being at least 18 years old and a Canadian citizen or permanent resident[@CES_2023].

The General Social Survey focuses on non-institutionalized individuals aged 15 and above residing in Canada's ten provinces.
Within each selected household, a random individual is chosen by an application to answer the survey following the initial household roster completion.
General Social Survey's questionnaire was developed through rigorous research and broad consultations with data users.
The Statistics Canada Questionnaire Design Resource Center (QDRC) conducted qualitative testing in four cities to identify effective questions and those requiring adjustments.
Based on the testing, QDRC prepared a comprehensive report with suggestions, which were integrated into the final survey design to enhance clarity and effectiveness.The survey employs a cross-sectional sample survey methodology, utilizing a combined frame of landline and cellular phone numbers from Census data, administrative records, and Statistics Canada's dwelling database.
This approach improves coverage of households linked to phone numbers.
The sampling strategy is stratified, with stratification occurring at the level of provinces and census metropolitan areas (CMA), and it relies on probability sampling techniques.
Information is collected from a single individual aged 15 or older per household, without the use of proxy responses.[@GSS_2022]

@@

# Discussion
Recent research highlights that the primary concerns for Canadian voters leading up to the 2025 election include the rising cost of living, housing, healthcare, the economy, and the environment. These issues are pivotal to the electorate, and the political parties’ stances on them are under intense scrutiny. Our analysis, drawn from five different polls, suggests a marginal preference for the Conservative Party over the Liberal Party. This trend is echoed in Barik's 2023 article, which indicates a public perception of the Conservative Party's superior handling of these key issues. Thus, our working hypothesis posits that the Conservative Party holds a slender advantage in the race for the upcoming election[@Barik_2023].

For our model development, we chose predictors from census and survey data, refined through thorough research and reflecting the public's primary concerns. We proceeded with **binary logistic regression models**—one for each major party—to predict voter preferences based on complex social backgrounds. After refining the models using Akaike's Information Criteria to eliminate superfluous predictors, we established age, province, religion, education, and pre-tax income as our final variables.

Post-stratification was applied to these predictors, revealing that the Conservative Party has a projected win probability of 25%, marginally outpacing the Liberal Party's 24.9%. These projections reinforce our initial hypothesis that the Conservative Party is the slight favorite for the 2025 election, based on our analysis of the data at hand. Our findings are in line with prior research, indicating that the main electoral contest will be between the Conservative and Liberal parties.

## Consideration of Potential Bias
One limitation we noted was the age of participants in the data from 2021. Participants under 18 then are of voting age now, in 2025. We opted to retain these young individuals in our dataset, hypothesizing that their burgeoning political views would mature by the election year. Nonetheless, the potential impact of their age on our prediction model merits further attention in subsequent studies.

## Pathways for Enhanced Predictive Accuracy
Our endeavor has shed light on the inherent complexities of electoral forecasting. The dynamic and deeply personal nature of voting behavior, compounded by ever-shifting political landscapes, renders the task formidable. For future projections, expanding the set of predictors could refine our model's accuracy, mitigating the risk of overlooking significant variables.

In conclusion, our binary logistic regression approach, supplemented with post-stratification, has yielded a data-driven forecast for the 2025 Canadian Election. Aligning with Barik’s analysis and multiple polls, our findings suggest the Conservative Party is poised for a narrow victory over the Liberal Party, informed by an intricate blend of sociopolitical factors and current Canadian sentiment[@Barik_2023].
